Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2021-09-21 mice_synthesizing/paper.Rmd

---
title: "Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*"
subtitle: "Anony*mice*d shareable data: Using *mice* to create multiply imputed synthetic datasets" 
author:
- Thom Volker, Gerko Vink & Stef van Buuren
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
bibliography: federated_imp.bib
csl: "apa-6th-edition.csl"
---

@@

- [ ] Explain FCS in short.
- [ ] Demonstrate procedure with `where-matrix` and explain rationale.
- [ ] Demonstrate that this is a reasonable approach, based on a figure that shows the observed and synthetic relationships side-by-side (e.g. `wgt ~ age + gen`).
- [ ] Observed values can be regarded as starting values for the imputation model. Since the observed data is used for creating the imputation models, no iterations are required. That is, the model is fixed, and not dependent on the imputed values.

@@

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to distribute research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Altough these regulations are much needed, privacy constraints are also ranked among the toughest challenges to overcome in the advancement of computational social science [@lazer_life_2009].

@@

@@

The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Accordingly, the complexity parameter is fixed at `cp = 1e-08` rather than the default value `1e-08` and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. By allowing for more complexity in the imputation model, bias in the estimates from the synthetic dataset is reduced. The predictor matrix was specified so that BMI was not imputed from a model, but calculated from the imputed height and weight values ($BMI = kg / m^2$), an approach called passive imputation. Additionally, BMI is excluded as a predictor for weight and height, because BMI values are directly dependent on the values of weight and height. All other variables $Y_{-j}$ are used as predictors for variable $Y_j$. Additionally, the number of imputations is set to $m = 5$.

## Synthe*miz*ing in practice

Discuss CART (read Reiter)
Add text about additional CART parameters cp and minbucket

@@

-- -- SINCE PARTIALLY SYNTHETIC DATA POOLING RULES DO NOT REQUIRE ADJUSTED VARIANCE ESTIMATES -- --

-- -- CAN WE REGARD THE OBSERVED VALUES AS THE STARTING VALUES FOR THE IMPUTATIONS ? -- --

<!-- The R-package `mice` [@mice] implements the fully conditional specification approach for missing data. Since the imputation approaches for missing data and synthetic data do not differ in terms of the imputation process, `mice` can be used for creating synthetic data as well, which will be shown in more detail in the remainder of this paper. In the next section, the pooling rules that are required to obtain inferences from the synthetic datasets are discussed. Thereafter, it will be investigated which variance estimator yields the best, that is, confidence valid, results in combination with the `mice` synthesizing approach. Also, it is shown that iterating over the conditional distributions in order to establish convergence is not required, due to the fact that new values are drawn directly from posterior distribution given the observed data. -->

>>>>

---
title: "Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*"
subtitle: "Anony*mice*d shareable data: Using *mice* to create multiply imputed synthetic datasets" 
author:
- Thom Volker & Gerko Vink
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
bibliography: federated_imp.bib
csl: "apa-6th-edition.csl"
---

@@

- [x] Explain FCS in short.
- [ ] Demonstrate procedure with `where-matrix` and explain rationale.
- [ ] Demonstrate that this is a reasonable approach, based on a figure that shows the observed and synthetic relationships side-by-side (e.g. `wgt ~ age + gen`).
- [ ] Observed values can be regarded as starting values for the imputation model. Since the observed data is used for creating the imputation models, no iterations are required. That is, the model is fixed, and not dependent on the imputed values.

@@

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to distribute research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Although these regulations are much needed, privacy constraints are also ranked among the toughest challenges to overcome in the advancement of computational social science [@lazer_life_2009].

@@

[[[MOVE UPCOMING TWO PARAGRAPHS TO SIMULATION SECTION]]]

@@

The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Accordingly, the complexity parameter is fixed at `cp = 1e-16 [[[CHECK THIS]]]` rather than the default value `1e-04` and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. By allowing for more complexity in the imputation model, bias in the estimates from the synthetic dataset is reduced. The predictor matrix was specified so that BMI was not imputed from a model, but calculated from the imputed height and weight values ($BMI = kg / m^2$), an approach called passive imputation. Additionally, BMI is excluded as a predictor for weight and height, because BMI values are directly dependent on the values of weight and height. All other variables $Y_{-j}$ are used as predictors for variable $Y_j$. Additionally, the number of imputations is set to $m = 5$.

# Synthesizing data with `mice`

To display the applicability of `mice` for generating synthetic data, we will display three distinct simulation set-ups that build up in complexity. The first case considers a multivariable regression model, in which the normally distributed outcome variable $y$ is regressed on three normally distributed predictor variables $\{x_1, x_2, x_3\}$, which are bivariately correlated with $\rho_{x_i, x_j} = 0.3$. The regression coefficients are specified in such a way that $\beta_2$ equals $2\beta_1$ and $\beta_3$ equals $3\beta_1$. Additionally, the proportion explained variance in $y$ equals $R^2=0.25$. Hence, the population regression weights equal $\beta_1 = 0.11$, $\beta_2 = 0.22$ and $\beta_3 = 0.33$.

The second simulation involves a logistic regression model, in which $y$ follows a Bernoulli distribution with probability 
$$
\frac{1}{1 + \text{exp}\{-X\beta\}},
$$
with predictor matrix $X$, consisting of 4 predictors. Here, $x_3$ moderates the relationship between $x_2$ and $y$, while $x_4$ is transformed into a categorical variable with three levels.

@@
